Arhangelskii's \(\alpha\)-principles and selection games

### 2019-11-05 articles/selection-alpha-games/article.tex
fix 1.5 to 1.1

\begin{abstract}
Arhangelskii's convergence properties \(\alpha_2\) and \(\alpha_4\)
may be characterized in terms of Scheeper's selection games.
We generalize these folklore results to hold for more general collections.
\end{abstract}

@@

\section{Main Results}

@@

@@

@@

@@

@@

Noting that \(\plI\win G_1(\Gamma_X,\Gamma_X)\) if and only if
\(\plI\win G_{fin}(\Gamma_X,\Gamma_X)\) [cite Kocinac], but
the same is not true of \(G_\star(\Gamma_{X,x},\Gamma_{X,x})\)
(i.e. there are \(\alpha_4\) spaces that are not \(\alpha_2\)
[cite Arhangelskii]),
we also ask the following.

\begin{question}
Is there an elegant condition on \(\mc A,\mc B\) guaranteeing
\(\plI\win G_1(\mc A,\mc B)\Rightarrow\plI\win G_{fin}(\mc A,\mc B)\)?
\end{question}

>>>>

\begin{abstract}
Arhangelskii's properties \(\alpha_2\) and \(\alpha_4\)
defined for convergent sequences
may be characterized in terms of Scheeper's selection principles.
We generalize these results to hold for more general collections
and consider these results in terms of selection games.
\end{abstract}

@@

@@

\section{On \(\alpha_2(\mc A,\mc B)\) and \(G_1(\mc A,\mc B)\)}

@@

While \(\alpha_2(\mc A,\mc B)\) involves infinite intersection and
\(G_1(\mc A,\mc B)\) involves single selections, the previous result is made
more intuitive given the following result, shown for \(\mc A=\mc B=\Gamma_{X,x}\)
by Nogura in [cite product of alpha spaces].

\begin{definition}
\(\alpha_2'(\mc A,\mc B)\) is the following claim:
if \(A_n\in\mc A\) for all \(n<\omega\), then there exists \(B\in\mc B\) such that
\(A_n\cap B\) is nonempty for all \(n<\omega\).
\end{definition}

(Note that \(\alpha_5\) is sometimes used in the literature in place of \(\alpha_2'\).)

\begin{proposition}
If \(\mc A\) is almost-\(\Gamma\)-like, then
\(\alpha_2(\mc A,\mc B)\) is equivalent to \(\alpha_2'(\mc A,\mc B)\).
\end{proposition}

\begin{proof}
The forward implication is immediate, so we assume \(\alpha_2'(\mc A,\mc B)\).
Given \(A_n\in\mc A\), we apply the almost-\(\Gamma\)-like property to obtain
\(A_n'=\{a_{n,m}:m<\omega\}\subseteq A_n\) such that
\(A_{n,m}=A_n\setminus\{a_{i,j}:i,j<m\}\in\mc A\) for all \(m<\omega\).

By applying \(\alpha_2'(\mc A,\mc B)\) to \(A_{n,m}\), we obtain
\(B\in\mc B\) such that \(A_{n,m}\cap B\) is nonempty for all \(n,m<\omega\).
Since it follows that \(A_n\cap B\) is infinite for all \(n<\omega\),
we have established \(\alpha_2(\mc A,\mc B)\).
\end{proof}

\section{On \(\alpha_4(\mc A,\mc B)\) and \(G_{fin}(\mc A,\mc B)\)}

A similar correspondence exists between \(\alpha_4(\mc A,\mc B)\)
and \(G_{fin}(\mc A,\mc B)\).

@@

\section{Perfect information and predetermined strategies}

@@

\section{Disjoint selections}

In each \(\alpha_i(\mc A,\mc B)\) principle, it is not required for the collection
\(\{A_n:n<\omega\}\) to be pairwise disjoint. However, in many cases it may as
well be.

\begin{definition}
For \(i\in\{1,2,3,4\}\) let \(\alpha_{i.1}(\mc A,\mc B)\) denote the claim that
\(\alpha_i(\mc A,\mc B)\) holds provided the collection \(\{A_n:n<\omega\}\)
is pairwise disjoint.
\end{definition}

Of course, \(\alpha_i(\mc A,\mc B)\) implies \(\alpha_{i.1}(\mc A,\mc B)\).
It's also immediate that \(\alpha_{i.1}(\mc A,\mc B)\) implies
\(\alpha_{i.1+1}(\mc A,\mc B)\) for the same reason that \(\alpha_i(\mc A,\mc B)\)
implies \(\alpha_{i+1}(\mc A,\mc B)\).

%Nyikos considered these half-variants
%in [cite Nyikos 1992] for \(\mc A=\mc B=\Gamma_{X,x}\), but his arguments
%can be generalized provided \(\mc A\) is \(\Gamma\)-like.

We take advantage of the following lemma.

\begin{lemma}[Lemma 1.2 of (cite Nyikos 92)]
Given a family \(\{A_n:n<\omega\}\) of infinite sets, there exist infinite subsets
\(A_n'\subseteq A_n\) such that \(\{A_n':n<\omega\}\) is pairwise disjoint.
\end{lemma}

\begin{proposition}
Let \(\mc A\) be \(\Gamma\)-like.
For \(i\in\{2,3,4\}\), \(\alpha_i(\mc A,\mc B)\) is equivalent to
\(\alpha_{i.1}(\mc A,\mc B)\).
\end{proposition}

\begin{proof}
Assume \(\alpha_{i.1}(\mc A,\mc B)\).
Let \(A_n\in\mc A\). By applying the previous lemma, we have
\(\{A_n':n<\omega\}\) pairwise disjoint with each \(A_n'\) being
an infinite subset of \(A_n\). Since \(\mc A\) is \(\Gamma\)-like,
\(A_n'\in\mc A\), so we have a witness \(B\in\mc B\) such that
\(A_n'\cap B\) satisfies \(\alpha_{i.1}(\mc A,\mc B)\) for all \(n<\omega\).
Since \(A_n'\subseteq A_n\), it follows that \(A_n\cap B\) satisfies
\(\alpha_{i}(\mc A,\mc B)\) for all \(n<\omega\).
\end{proof}

It's also true that \(\alpha_1(\Gamma_{X,x},\Gamma_{X,x})\)
is equivalent to \(\alpha_{1.1}(\Gamma_{X,x},\Gamma_{X,x})\),
which is captured by the following theorem.

\begin{proposition}
Let \(\mc A\) be a \(\Gamma\)-like collection closed under finite unions.
Then \(\alpha_1(\mc A,\mc A)\) is equivalent to
\(\alpha_{1.1}(\mc A,\mc A)\).
\end{proposition}

\begin{proof}
Let \(A_n\in\mc A\) and assume \(\alpha_{1.1}(\mc A,\mc A)\).
To apply the assumption, we will define a pairwise disjoint
collection \(\{A_n':n<\omega\}\). First let \(0'=0\) and \(A_0'=A_0\).
Then suppose \(m'\geq m\) and \(A_m'\subseteq A_{m'}\subseteq\bigcup_{i\leq m}A_i'\) 
are defined for all \(m\leq n\).

If \(A_k\setminus\bigcup_{m\leq n}A_m'\) is finite for \(k>n'\), let
\(B=\bigcup_{m\leq n'}A_m\in\mc A\). This \(B\) then witnesses 
\(\alpha_1(\mc A,\mc A)\) since \(A_k\setminus B\) is finite for all \(k<\omega\).

Otherwise pick the minimal \((n+1)'>n\) where 
\(A_{n+1}'=A_{(n+1)'}\setminus\bigcup_{m\leq n}A_m'\) is infinite.
It follows that \(A_{n+1}'\subseteq A_{(n+1)'}\subseteq \bigcup_{m\leq n+1}A_m'\).
By construction, \(\{A_n':n<\omega\}\) is a pairwise disjoint collection of
members of \(\mc A\), and we may
apply \(\alpha_{1.1}(\mc A,\mc A)\) to obtain \(B\in\mc A\) where
\(A_n'\setminus B\) is finite for all \(n<\omega\).

Finally let \(k<\omega\). If \(k=n'\) for some \(n<\omega\), then 
\(A_k\setminus B=A_{n'}\setminus B\subseteq (\bigcup_{m\leq n}A_m')\setminus B\) 
is finite.
Otherwise, \(n'<k<(n+1)'\) for some \(n<\omega\).
Then 
\((A_k\setminus\bigcup_{m\leq n}A_m')\setminus B\subseteq A_k\setminus\bigcup_{m\leq n}A_m'\)
is finite, and 
\((A_k\cap\bigcup_{m\leq n}A_m')\setminus B\subseteq(\bigcup_{m\leq n}A_m')\setminus B\)
is finite, showing \(A_k\setminus B\) is finite.
\end{proof}

TODO add discussion of \(\alpha_{1.5}\) here.

For this paragraph we adopt the conventional assumption that
\(\Gamma_{X,x}\) is restricted to countable sets.
Nyikos showed a consistent example
where \(\alpha_2(\Gamma_{X,x},\Gamma_{X,x})\) fails to imply
\(\alpha_{1.5}(\Gamma_{X,x},\Gamma_{X,x})\), and a consistent example where
\(\alpha_{1.5}(\Gamma_{X,x},\Gamma_{X,x})\) fails to imply
\(\alpha_{1}(\Gamma_{X,x},\Gamma_{X,x})\) [cite Nyikos 92].
On the other hand, Dow showed that \(\alpha_2(\Gamma_{X,x},\Gamma_{X,x})\)
implies \(\alpha_{1}(\Gamma_{X,x},\Gamma_{X,x})\) in the Laver model
for the Borel conjecture [cite Dow 1990]; the author conjectures
that this model (specifically, the fact that
every \(\omega\)-splitting family contains an \(\omega\)-splitting
family of size less than \(\mathfrak b\) in this model) witnesses
an affirmative answer to the following question.

\begin{definition}
A \(\Gamma\)-like collection is \term{strongly-\(\Gamma\)-like}
if the collection is closed under finite unions and each member
is countable.
\end{definition}

\begin{question}
Let \(\mc A,\mc B\) be strongly-\(\Gamma\)-like
and \(\mc B\subseteq\mc A\).
Is it consistent that \(\alpha_2(\mc A,\mc B)\)
implies \(\alpha_1(\mc A,\mc B)\)?
\end{question}

@@

Noting that \(\plI\win G_1(\Gamma_X,\Gamma_X)\) if and only if
\(\plI\win G_{fin}(\Gamma_X,\Gamma_X)\) [cite Kocinac], but
the same is not true of \(G_\star(\Gamma_{X,x},\Gamma_{X,x})\)
(i.e. there are \(\alpha_4\) spaces that are not \(\alpha_2\)
[cite Shakhmatov convergence algebraic structure]),
we also ask the following.

\begin{question}
Is there a natural condition on \(\mc A,\mc B\) guaranteeing
\(\plI\win G_1(\mc A,\mc B)\Rightarrow\plI\win G_{fin}(\mc A,\mc B)\)?
\end{question}

\section{Acknowledgements}

The author would like to thank Dr. Alan Dow and Dr. Jared Holshouser
for their discussions on the subject of the paper.
